---
title: "EcostarWP4_Platform Clustering"
author: "Clement Garcia"
date: "2024-03-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Installing required libraries and loading them:
pkgs = c("tidyverse", "vegan", "patchwork", "devtools", "PCDimension", "ggrepel", "rnaturalearth", "sinkr")
for(p in pkgs){
  if(!require(p, character.only = TRUE)) install.packages(p)
  library(p, character.only = TRUE)
} 
install_github("marchtaylor/sinkr")

## Data
setwd("C:/Users/cg05/OneDrive - CEFAS/Science/Project - Commercial/INSITE/INSITE II/online/EcostarWP4/")
load("./data_processed/NS_UKBandENV.RData")
load("./data_processed/UKBenthos_spePlat.RData")
NS_Plat_location<-read.csv("./data_processed/NS_Plat_location.csv")

```

We ordinate the different platforms according to their environmental variables (from the DataPre script) and then cluster to control for the environmental context when testing for difference in structure and function close and far from the platforms.
Depth: Water depth (in m)
MD0: Median grain size of sediment (Phi units).
Mud: Silt/clay content of the sediment (percentage by weight of the sub 63 μm fraction of the sediment).
OM: Organic content of the sediment (percentage by weight of organic material in the sediment).
Distance: Distance to coast (in m)
Current: Bottom current

## PCA

```{r PCAprep, echo=FALSE}
## Data reconstruction
#Formatting to have a numeric data frame to scale
rownames(NS_ENV)<-NS_ENV$PLAT_CAT
NS_ENV$PLAT_CAT<-NULL
NS_ENVsc<-scale(NS_ENV)

#Completing the dataset because of missing value using dineof function (from sinkr)
set.seed(3180)
temp<-dineof(NS_ENVsc)
NS_ENVsc<-temp$Xa

```
```{r PCA, echo=TRUE}
## PCA
PLAT.PCA<-rda(NS_ENVsc)
## Extracting the eigenvalues
length(evPLAT <- PLAT.PCA$CA$eig)

## Broken stick method helps to justify how many axes to select
bsPLAT<-brokenStick(1:6, 6)
eigPLAT<<-data.frame(EIGEN = 1:6, eigP = (PLAT.PCA$CA$eig/sum(PLAT.PCA$CA$eig))*100, bsPLAT = bsPLAT/sum(bsPLAT)*100)

ggplot(eigPLAT, aes(x=EIGEN, y=eigP)) +
  geom_bar(stat="identity", aes(fill = eigP>15))+
  geom_line(aes(y=bsPLAT), colour = "red")+
  geom_point(aes(y=bsPLAT), colour = "red")+
  scale_fill_grey(start = 0.8, end = 0, guide = 'none')+
  labs(x = 'PC', y = '% Variance')+
  theme_bw()+
  ggtitle("Platform - Environment")

## PCA parameters*****************************
axes <- c(1,2) # scree plot 2 axes are enough
site.scaling <- 1 # Default
env.scaling <- 2 # Default
goodness.axis <- 2 # Number of axis kept
goodness.tresh <- 2/6 # Ease of representation
#*********************************************

## c_sq data
PLAT.scores <- as.data.frame(scores(PLAT.PCA, scaling = site.scaling, display="sites", choices=axes))
axes_name.PLAT <- colnames(PLAT.scores)

```
Broken stick shows that 3 principal axes are necessary but for clarity we only kept 2.
The following show that the k-mean method with and SSI criterion to justify how many clusters the different platforms belong to. Results show that ideally there would be 14 or 8 clusters but that would be difficult to manipulate.
We therefore only kept 5 for clarity.

```{r Cluster, echo=TRUE}
## Cluster analysis on the PCA output
set.seed(8110)

## SSI method to justify the number of cluster
PLAT.PCA.KM<-cascadeKM(PLAT.scores, inf.gr=2, sup.gr=20, iter=100, criterion = "ssi")
summary(PLAT.PCA.KM)
plot(PLAT.PCA.KM, sortg=T)
## Number of cluster 5, 8 & 14
g4<-as.data.frame(PLAT.PCA.KM$partition[,4])
g7<-as.data.frame(PLAT.PCA.KM$partition[,7])
g13<-as.data.frame(PLAT.PCA.KM$partition[,13])

## number of instances
xtabs( ~g4[,1],g4) 
xtabs( ~g7[,1],g7) 
xtabs( ~g13[,1],g13) 

## Add clusters to PC scores
PLAT.scores<- cbind(PLAT.scores, g4, g7, g13)
colnames(PLAT.scores)[c(3:5)]<-c("cluster5", "cluster8", "cluster14")

## Add clusters to raw data 
NS_ENVsc<-as.data.frame(NS_ENVsc)
NS_ENVsc$PLAT_CAT<-rownames(NS_ENVsc)
NS_ENVsc<-cbind(NS_ENVsc, PLAT.scores)

PLATdf<-NS_ENVsc %>% 
  gather(environment, value, Depth:Current)

gPLAT<-ggplot(PLATdf, aes(x=environment, y=value, colour = factor(cluster5))) + 
  geom_boxplot() +
  stat_summary(fun=mean, geom="point", shape=5, size=4)+
  theme_bw() +
  xlab("")+
  scale_colour_discrete(name="Cluster")+
  theme(axis.text.x  = element_text(angle=45, vjust=0.5, size=9))+
  facet_grid(.~cluster5)

gPLAT
#ggsave('./figures/ClusPlat.png')
```
The boxplot show the characteristics of each of the 5 clusters.
Cluster 1 is intermediate muddy
Cluster 2 is deep muddy
Cluster 3 is the average cluster with no discernible characteristics
Cluster 4 is shallow muddy
Cluster 5 is shallow coarse
```{r PCAoutput, echo=TRUE}
## Descriptor scores
env.scores <- scores(PLAT.PCA, scaling=env.scaling, display="species", choices=axes)

## Retrieve the goodness of fit of the species
env_fit.sum <- goodness(PLAT.PCA, model="CA",statistic="explained")

## Combine scores and  goodness of fit for the selected axis
env.scores <- data.frame(environment=rownames(env.scores), env.scores, fit=env_fit.sum[,goodness.axis])

## Calculate the variance represented by each axis
var_axes <- round(PLAT.PCA$CA$eig/sum(PLAT.PCA$CA$eig)*100,2)

## Representativity of species on the three axes
AVrep<-ggplot(env.scores, aes(x= reorder(environment, -fit), y= fit)) + 
  geom_bar(stat="identity", aes(fill = fit>0.33)) +
  geom_hline(yintercept = 0.33)+
  scale_fill_grey(start = 0.8, end = 0, guide = 'none')+
  scale_x_discrete(name="Representativity") +
  scale_y_continuous(name="Taxa")+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=45, vjust=1, hjust=1, size=10))+
  ggtitle("Goodness")


## c_sq plot (scaling 1)
## PC1 v PC2
pPLAT.PCA<-ggplot(data=PLAT.scores, 
                aes(x = PC1, y = PC2, colour = factor(cluster5)))+
  geom_point(size=1)+
  geom_hline(yintercept=0, linetype="dotted") +
  geom_vline(xintercept=0, linetype="dotted") +
  stat_ellipse(type="norm", geom="polygon", alpha=0.1, linetype=2)+
  xlab(paste(axes_name.PLAT[1],": ",var_axes[1],"%"))+
  ylab(paste(axes_name.PLAT[2],": ",var_axes[2],"%"))+
  theme_bw()+
  theme(legend.position="FALSE")+
  ggtitle("Platform cluster")

## Environment plot (scaling 2)
pPLAT.PCA.Env<-ggplot()+
  geom_point(data=PLAT.scores, 
             aes(x = PC1*2, y = PC2*2, colour = as.factor(cluster5)),
             size=2, alpha=0.3)+
  geom_segment(data=env.scores,
               aes(x=0, xend=PC1, y=0, yend=PC2),
               color="black", arrow=arrow(length=unit(0.01,"npc")))+
  geom_text_repel(data=env.scores,
                  aes(x=PC1,y=PC2,
                      label=env.scores[, "environment"]),
                  color="black", fontface='bold', size=3)+
  geom_hline(yintercept=0, linetype="dotted") +
  geom_vline(xintercept=0, linetype="dotted") +
  xlab(paste(axes_name.PLAT[1],": ",var_axes[1],"%"))+
  ylab(paste(axes_name.PLAT[2],": ",var_axes[2],"%"))+
  theme_bw()+
  theme(axis.title = element_text(size=10),
        axis.text  = element_text(size=10),
        legend.position="FALSE")

#pdf('./figures/PCA_PLAT.pdf',
#    width = 10, height = 6) 
pPLAT.PCA.Env + AVrep + plot_annotation(tag_levels = 'a')
#dev.off()

PLAT.scores$PLAT_CAT<-rownames(PLAT.scores)
PLAT.scores<-left_join(PLAT.scores, NS_Plat_location)
world <- ne_countries(scale = "medium", returnclass = "sf")


ggplot()+
  geom_sf(data = world)+
  geom_point(data = PLAT.scores,
             aes(x = LONG_PLAT_avg, y=LAT_PLAT_avg, colour = factor(cluster5)), size=0.5)+
  theme(legend.position="none")+
  xlab('')+
  ylab('')+
  ylim(48, 63)+
  xlim(-10, 8)
#ggsave('./figures/MapPlat.png')

```
Geographic representation show that Cluster 1 (red - intermediate muddy), 2 (yellow - deep muddy) and 4 (blue - shallow muddy) ranges from the East of the North North Sea by the Norwegian trench (2), East of Scotland (1) and south of the North North Sea (3). The two remaining clusters are the average cluster (3) which is spread equally across the central North Sea from North to South and the 5 (purple) is characterised by the shallow coarse area in the south North Sea.

```{r wrapUp, echo=FALSE}
## Cluster number to station data
PLAT.Clus<-PLAT.scores[, c("cluster5", "cluster8", "cluster14", "PLAT_CAT")]
UKB_Station<-left_join(UKB_Station, PLAT.Clus)

## Formatting station file
colnames(UKB_Station)<-c("site", "survey", "year", "stat_no", "plat_name", "plat_no", "lat_stat", "long_stat", "lat_plat", "long_plat", "lat_plat_dec", "long_plat_dec", "lat_stat_dec", "long_stat_dec", "distance", "bearing", "depth_stat", "depth_plat", "plat_cat", "cluster5", "cluster8", "cluster14")

STAT.clus<-UKB_Station[, c("site", "distance", "cluster5", "cluster8", "cluster14")]

spePlat<-left_join(spePlat, STAT.clus)
spePlat$B<-spePlat$A * spePlat$Meanbodymass
#save(spePlat, file="./spePlat.RData")

```

## END